---
title: "Computing gene signatures on T Lymphocytes"
author: "Laura Jiménez Gracia"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.align = "center")
```

# Introduction

# Pre-processing
## Libraries
```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(Seurat)
library(UCell)
library(pheatmap)
```

## Parameters
Here we will define and load objects and functions that will be used throughout the document.
```{r}
# Paths
path_project_metadata <- here::here("01_cellranger_mapping/data/FIXnCUT_metadata.csv")
path_r_objects_in <- here::here("03_clustering_annotation/results/R_objects")
path_r_results <- here::here("04_GEX_analysis/gene_signatures/results")

# Pseudobulk function (equivalent to DecoupleR)
get_pseudobulk <- function(seurat_obj, sample_col = "individual", groups_col = "cell_type", min_cells = 0, min_counts = 0) {
  # Extract raw integer counts
  counts_matrix <- GetAssayData(seurat_obj, slot = "counts")
  # Extract metadata
  metadata <- seurat_obj@meta.data
  
  # Create a combined group identifier
  metadata$group_id <- paste(metadata[[sample_col]], metadata[[groups_col]], sep = "_")
  
  # Group cells by the combined identifier
  cell_groups <- split(rownames(metadata), metadata$group_id)
  # Aggregate counts by summing across all cells in each group
  pseudobulk_counts <- sapply(cell_groups, function(cells) {
    if (length(cells) >= min_cells) {
      summed_counts <- Matrix::rowSums(counts_matrix[, cells, drop = FALSE])
      if (sum(summed_counts) >= min_counts) {
        return(summed_counts)
      }
    }
    return(rep(0, nrow(counts_matrix)))
  })
  # Convert to matrix format
  pseudobulk_counts <- as.matrix(pseudobulk_counts)
  # Remove empty columns
  valid_cols <- colSums(pseudobulk_counts) > 0
  pseudobulk_counts <- pseudobulk_counts[, valid_cols, drop = FALSE]
  
  # Create sample-level metadata
  sample_metadata <- unique(metadata[, c(sample_col, groups_col), drop = FALSE])
  sample_metadata$group_id <- paste(sample_metadata[[sample_col]], sample_metadata[[groups_col]], sep = "_")
  rownames(sample_metadata) <- sample_metadata$group_id
  # Filter metadata to match valid columns
  sample_metadata <- sample_metadata[valid_cols, , drop = FALSE]
  
  # Create a new Seurat object with pseudobulk data
  pseudobulk_seurat <- CreateSeuratObject(
    counts = pseudobulk_counts,
    meta.data = sample_metadata
  )
  
  return(pseudobulk_seurat)
}
```

```{r}
# Load data
seurat_obj <- readRDS(paste0(path_r_objects_in, "/VEIGAEST_clustering_annotation_full_cleaned.rds"))
seurat_obj
```

# Pseudobulking across T Lymphocytes
```{r}
# Subsetting on T Lymphocytes
seurat_obj_T <- subset(seurat_obj, subset = cell_lineages == "T Lymphocytes")
seurat_obj_T
rm(seurat_obj)

seurat_obj_T$cell_types <- as.factor(as.character(seurat_obj_T$cell_types))

# Pseudobulk
seurat_obj_T_pseudo <- get_pseudobulk(seurat_obj_T, sample_col = "treatment", groups_col = "cell_types", min_cells = 0, min_counts = 0)
seurat_obj_T_pseudo


Tcell_types_order <- c("Stem-like CD8 T cells",
                      "Effector CD8 T cells",
                      "Effector Memory CD8 T cells",
                      "IFN-responder CD8 T cells",
                      "Exhausted CD8 T cells",
                      "Cycling CD8 T cells", 
                      "Exhausted CD4 T cells",
                      "Tregs", "Cycling Tregs", 
                      "Th",
                      "MAIT", 
                      "g/dv2-17 cells",
                      "NKT cells"
                      )

seurat_obj_T_pseudo$cell_types <- factor(x = seurat_obj_T_pseudo$cell_types,
                                        levels = Tcell_types_order)

# Normalization
seurat_obj_T_pseudo <- NormalizeData(seurat_obj_T_pseudo, normalization.method = "LogNormalize", scale.factor = 10000)
```


# Computing gene signatures

## In-house gene signatures
```{r}
bulkRNAseq_gene_signatures <- list(
  "cell_cycle_arrest" = c("Cdkn1b", "Ccng2", "Cdkn2c", "Cdkn2d", "Cdkn1a", "Cdkn2b"),
  "cell_cycle_progression" = c("Mcm5", "Cdk4", "Mcm6", "Ccnb1", "Mcm2", "Mcm4", "E2f1", "Cdk2", "Mcm3", "Mcm7", 
                               "Ccnd2", "Cdk8", "Cdk11b", "Ccnd3", "Tfdp1", "Cdk16", "Cdk5", "Cdk19", "E2f2", "Ccnb2", "Cdk1"),
  
  "apoptosis" = c("Pmaip1", "Bak1", "Bax", "Bid", "Cflar", "Fas", "Bnip3", "Bcl2l11", "Bbc3"),
  "cell_survival" = c("Birc5", "Sirt3", "Traf5", "Bcl2a1d", "Bcl2a1b", "Id2", "Traf2", "Bcl2l1", "Hmox1", "Tnfrsf4", "Bcl2", "Tnfrsf9", "Traf3"),
  
  "stemLike_phenotype" = c("Bhlhe40", "Myb", "Cxcl10", "Il7r", "Ccr7", "Ikzf2", "Foxo1", "Slamf6", "Lef1",
                           "Bcl6", "Bach2", "Klf3", "Tcf7", "S1pr1", "Id3", "Klf2", "Sell"),
  "quiescence" = c("Myb", "Il7r", "Ccr7", "Ikzf2", "Foxo1", "Slamf6", "Lef1", "Bcl6", "Bach2", "Klf3", "Tcf7", "S1pr1", "Id3", "Klf2", "Sell"),
  "exhaustion" = c("Nfatc1", "Cd7", "Ctla4", "Cd101", "Pdcd1", "Nfil3", "Tox", "Xbp1", "Nr4a3",
                   "Nr4a1", "Nr4a2", "Cd244a", "Havcr2", "Entpd1", "Cd160", "Cd38"),
  
  "Tc1_phenotype" = c("Fasl", "Gzmb", "Il2ra", "Ccl3", "Tnf", "Ifng", "Icos", "Prf1", "Cd69", "Prdm1", "Irf4", "Ccl4", "Tnfrsf8", 
                      "Gzmg", "Gzmd", "Gzme", "Gzmf", "Tbx21", "Runx3", "Gzma", "Gzmk", "Eomes", "Batf", "Gzmc", "Itgal", "Cd44", "Ccl5"),
  
  "glucose_metabolism" = c("Hk2", "Pgm2", "Pdk3", "Eno3", "Bpgm", "Pdk2", "Ldhb", "Pgm2l1", "Pgam2", "Ldha", "Pfkp", 
                           "Eno1", "Eno1b", "Slc2a1", "Scl16a3", "Pgam1", "Tpi1", "Pkm", "Hif1a", "Pfkm", "Aldoa", "Pfkfb3", 
                           "Ipfk2", "Acss1", "Acss2", "Aldoc", "Gpi1", "Pgk1", "Pygl", "Slc2a3"),
  "mitochondrial_metabolism" = c("Slc25a23", "Ndufaf1", "Nrf1", "Ndufb9", "Atp5o", "Ndufa9", "Ndufs8", "Fabp5", "Atp5b", "Ndufa2",
                                 "Uqcr10", "Ndufa5", "Ptpmt1", "Ndufb7", "Uqcrq", "Ndufv1", "Cox6b2", "Ndufa4", "Cox5a", "Atp5k", 
                                 "Cox7b", "Atp5l", "Atp5g3", "Ndufa12", "Ndufs6", "Ndufb3", "Ndufa1", "Emc8", "Cox6a1", "Mtch2",
                                 "Sdhb", "Ndufaf2", "Ndufb6", "Ndufs1", "Atpaf2", "Atpif1", "Ndufaf4")
)
```


### Genes
```{r fig.width=10, fig.height=4}
for (gene_sig in names(bulkRNAseq_gene_signatures)) {
  gg_gene_sig <- DotPlot(seurat_obj_T_pseudo,
        features = bulkRNAseq_gene_signatures[[gene_sig]],
        group.by = "treatment",
        cols = "RdBu",
        dot.scale = 5
        ) + 
    labs(title = gene_sig) +
   theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
  print(gg_gene_sig)
}
```

```{r fig.width=16, fig.height=10}
for (gene_sig in names(bulkRNAseq_gene_signatures)) {
  gg_gene_sig <- DotPlot(seurat_obj_T_pseudo,
        features = bulkRNAseq_gene_signatures[[gene_sig]],
        group.by = "cell_types",
        split.by = "treatment",
        cols = "RdBu",
        dot.scale = 5
        ) + 
    coord_flip() +
    labs(title = gene_sig) +
   theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
  print(gg_gene_sig)
}
```

### Computing gene signature with UCell

The function AddModuleScore_UCell() allows operating directly on Seurat objects. UCell scores are calculated from raw counts or normalized data, and returned as metadata columns. The example below defines some simple signatures, and applies them on single-cell data stored in a Seurat object.

```{r fig.width=8, fig.height=5}
# UCELL GENE SIGNATURE SCORE
## Using positive and negative gene sets, GENE+ or GENE-
seurat_obj_T_pseudo <- AddModuleScore_UCell(seurat_obj_T_pseudo, 
                                   features=bulkRNAseq_gene_signatures, name=NULL)
head(seurat_obj_T_pseudo@meta.data)
```

```{r}
seurat_obj_T_pseudo_metadata <- seurat_obj_T_pseudo@meta.data
saveRDS(seurat_obj_T_pseudo_metadata, paste0(path_r_results, "/VEIGAEST_Tlymphocytes_GeneSignatureScores_metadata.rds"))
```

### Gene signatures
```{r}
gene_signatures_scores_df <- seurat_obj_T_pseudo@meta.data[, c("cell_types", "treatment", names(bulkRNAseq_gene_signatures))]

gene_signatures_scores_df$cell_types <- factor(gene_signatures_scores_df$cell_types, levels = Tcell_types_order)
gene_signatures_scores_df <- gene_signatures_scores_df %>%
  arrange(cell_types, treatment)

gene_signatures_scores_df$condition <- paste0(gene_signatures_scores_df$cell_types, "__", gene_signatures_scores_df$treatment)
gene_signatures_scores_df$cell_types <- NULL
gene_signatures_scores_df$treatment <- NULL
rownames(gene_signatures_scores_df) <- gene_signatures_scores_df$condition
gene_signatures_scores_df$condition <- NULL
gene_signatures_scores_df
```

```{r fig.height=6, fig.width=15}
pheatmap(t(gene_signatures_scores_df), 
         scale = "row",
         cluster_rows = FALSE, # Cluster rows
         cluster_cols = FALSE, # Cluster columns
         show_rownames = TRUE, # Show row names
         show_colnames = TRUE, # Show column names
         color = colorRampPalette(c("blue", "white", "red"))(100), # Color scale
         border_color = "white",
         main = "Bulk RNA-seq Gene Signatures")
```

## Li, J. et al. CancerCell (2023)
https://doi.org/10.1016/j.ccell.2023.04.011

```{r}
CRC_gene_signatures <- list(
  "effector_molec" = c("Ifng", "Nkg7", "Prf1", "Gzmk", "Gzma", "Faslg"),
  "costimulation_activation" = c("Cd27", "Cd82", "Tnfrsf4", "Tnfrsf9", "Cxcr3"),
  "coinhibitory_receptors" = c("Pdcd1", "Lag3", "Tigit", "Havcr2", "Ctla4", "Tnfrsf18"),
  "exhaustion" = c("Tox", "Entpd1", "Batf", "Prdm1"),
  "chemokines" = c("Ccl3", "Ccl4", "Ccl5", "Cxcl13")
)
```


### Genes
```{r fig.width=10, fig.height=3}
for (gene_sig in names(CRC_gene_signatures)) {
  gg_gene_sig <- DotPlot(seurat_obj_T_pseudo,
        features = CRC_gene_signatures[[gene_sig]],
        group.by = "treatment",
        cols = "RdBu",
        dot.scale = 5
        ) + 
    labs(title = gene_sig) +
   theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
  print(gg_gene_sig)
}
```

```{r fig.width=16, fig.height=7}
for (gene_sig in names(CRC_gene_signatures)) {
  gg_gene_sig <- DotPlot(seurat_obj_T_pseudo,
        features = CRC_gene_signatures[[gene_sig]],
        group.by = "cell_types",
        split.by = "treatment",
        cols = "RdBu",
        dot.scale = 5
        ) + 
    coord_flip() +
    labs(title = gene_sig) +
   theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
  print(gg_gene_sig)
}
```

### Computing gene signature with UCell

The function AddModuleScore_UCell() allows operating directly on Seurat objects. UCell scores are calculated from raw counts or normalized data, and returned as metadata columns. The example below defines some simple signatures, and applies them on single-cell data stored in a Seurat object.

```{r fig.width=8, fig.height=5}
# UCELL GENE SIGNATURE SCORE
## Using positive and negative gene sets, GENE+ or GENE-
seurat_obj_T_pseudo <- AddModuleScore_UCell(seurat_obj_T_pseudo, 
                                   features=CRC_gene_signatures, name=NULL)
head(seurat_obj_T_pseudo@meta.data)
```

### Gene signatures
```{r}
gene_signatures_scores_df <- seurat_obj_T_pseudo@meta.data[, c("cell_types", "treatment", names(CRC_gene_signatures))]

gene_signatures_scores_df$cell_types <- factor(gene_signatures_scores_df$cell_types, levels = Tcell_types_order)
gene_signatures_scores_df <- gene_signatures_scores_df %>%
  arrange(cell_types, treatment)

gene_signatures_scores_df$condition <- paste0(gene_signatures_scores_df$cell_types, "__", gene_signatures_scores_df$treatment)
gene_signatures_scores_df$cell_types <- NULL
gene_signatures_scores_df$treatment <- NULL
rownames(gene_signatures_scores_df) <- gene_signatures_scores_df$condition
gene_signatures_scores_df$condition <- NULL
gene_signatures_scores_df
```

```{r fig.height=6, fig.width=15}
pheatmap(t(gene_signatures_scores_df), 
         scale = "row",
         cluster_rows = FALSE, # Cluster rows
         cluster_cols = FALSE, # Cluster columns
         show_rownames = TRUE, # Show row names
         show_colnames = TRUE, # Show column names
         color = colorRampPalette(c("blue", "white", "red"))(100), # Color scale
         border_color = "white",
         main = "Li, J. et al. CancerCell (2023) Gene Signatures")
```

## Krishna, S., et al. Science (2020)

```{r}
HumanCancer_gene_signatures <- list(
"Toxsig_scott" = c("Igkc", "Avil", "Tox", "Klrb1c", "Megf11", "Scd1", "Rp23-284k1.6",
  "Pif1", "Tnfrsf8", "Tmem163", "Nr2f6", "Serpinb6b", "Fgr", "Lrrk1", "Suv39h2", "Esm1",
  "Cenpe", "Ccr2", "Klrc1", "Nt5dc2", "Cd244", "Aspm", "Fam20a", "Zdhhc2", "Ubash3b",
  "Sapcd2", "Scd2", "Kif20b", "Ccr5", "Acsl3", "Wee1", "Hmmr", "Sh2d5", "Myo1e", "Cenpf",
  "Havcr2", "Eomes", "Pak6", "Ahr", "Arsb", "Klrk1", "Il10", "Ckap5", "Alyref", "Gen1",
  "Il24", "Ect2", "Mki67", "Gtf2ird1", "Gzma", "Hmgcs1", "Bspry", "Zbtb32", "Tubb6",
  "Bub1", "Anln", "Sgol2a", "Ulbp1", "Lmnb1", "Ccna2", "Cobll1", "Hras", "Stk39", "Fancb",
  "Srsf2", "Mpp6", "Fhl2", "Kif11", "Ncapg2", "Fkbp5", "Ankle1", "Stil", "Prc1", "Gpr141",
  "Mtmr1", "Nfil3", "Bub1b", "Itgav", "Swap70", "Usp1", "Entpd1", "Rac1", "Sgol1",
  "Ncapg", "Pgp", "Spag5", "Slc25a13", "Emilin2", "Trim59", "Ccnb1", "Mastl", "Asb2",
  "Srsf9", "Cdc25c", "Man2a1", "Kntc1", "Aurka", "Topbp1", "Cep55", "Ugcg"),

"exhaust_1_feldman" = c("Spc25", "Cdca5", "Esco2", "Cdc45", "Zwint", "Shcbp1", "Dlgap5", "Rad51", "Kif18b", "Rrm2",
  "Birc5", "Tk1", "Hjurp", "Ube2c", "Ccnb2", "Cenpw", "Gins2", "Rad51ap1", "Dtl", "Spc24",
  "Cdca3", "Pkmyt1", "Melk", "Anln", "Cdca8", "Kiaa0101", "Ggh", "AurkB", "Asf1b", "Cdc20",
  "Ncapg", "Dhfr", "Kifc1", "Tyms", "Ckap2l", "Clspn", "Mlf1ip", "Troap", "Kif2c", "Wdr34",
  "Cdk1", "Kif23", "Plk1", "Top2a", "Nuf2", "Hmgb3", "Aspm", "Mcm2", "Orc6", "Casc5",
  "Cenph", "Fen1", "Brca1", "Mcm4", "Timeless", "Mki67", "Cdkn3", "Apobec3b", "Ccnb1", "Tpx2",
  "Ncapg2", "Kif11", "Tcf19", "Ube2t", "Spag5", "Brca2", "Ccna2", "Bub1b", "Chek1", "Bub1",
  "Fanci", "Cenpm", "Rnaseh2a", "Hirip3", "Mad2l1", "Ccnf", "Stmn1", "Smc2", "Cks1b", "Paics",
  "NcapH", "Atad5", "Prc1", "Rfc5", "Cenpf", "Cenpn", "Cdca7", "Chtf18", "Cenpe", "Wdr76",
  "Fbxo5", "Cdca7l", "Rfc4", "Pold1", "Lrr1", "Racgap1", "Snrnp25", "Kntc1", "Nudt1", "Acot7",
  "Cdk2", "Gmnn", "Tmem106c", "Lig1", "Fancg", "Mxd3", "Pcna", "C19orf48", "Pola2", "Pole",
  "Ndc80", "Mcm3", "Cdk4", "Slc43a3", "Tubg1", "Nme1", "Haus8", "Mcm7", "Ncapd2", "Rfc2",
  "Shmt1", "DtymK", "Phf19", "Pold3", "Dctpp1", "Pop7", "Ahcy", "Tex30", "Nusap1", "Dut",
  "Kif18a", "Mrpl17", "Ggct", "Cenpk", "Haus4", "Pold2", "Lmnb1", "Blm", "Prim1", "Mt1e",
  "Cdca4", "Rrm1", "Rbbp8", "Ncapd3", "Tfdp1", "Ung", "Atad2", "Acat2", "Hadh", "Pafah1b3",
  "Yeats4", "Rpa3", "Timm10", "Mcm5", "Fancd2", "Rad51c", "Icmt", "Hist1h1d", "Lsm5", "Ssrp1",
  "Hist1h4c", "Dpysl2", "Ttf2", "Eef1e1", "Nup37", "Ebna1bp2", "Ccdc167", "Msh2", "Fh", "Ddx11",
  "Mrpl12", "Prdx4", "Dnajc9", "Anapc15", "Itgb3bp", "Magohb", "Mms22l", "Actl6a", "Bola3", "Usp1",
  "Nabp2", "Mcm6", "Ruvbl2", "Vrk1", "NcapH2", "Hells", "Kif22", "Whsc1", "Myl6b", "Pdcd5",
  "Shmt2", "Hist1h1c", "Ranbp1", "Sccpdh", "Haus1", "Mrps23", "Ubr7", "Polr2h"),

"exhaust_2_feldman" = c("Gem", "Layn", "Vcam1", "Rdh10", "Fam3c", "Kir2dl4", "Tnfrsf18", "Mtss1", "Cadm1", "Entpd1",
  "Etv1", "Afap1l2", "Tnfrsf9", "Nab1", "Peli1", "Dfnb31", "Ctla4", "Hspb1", "Fkbp4", "Namptl",
  "Myo7a", "Cxcl13", "Golim4", "Phlda1", "Dnaja4", "Tgif1", "Havcr2", "Aplp2", "Gpr56", "Bpgm",
  "Sec14l1", "Tnip3", "Metrnl", "Hsph1", "Klrc2", "Pmaip1", "Dusp4", "Igflr1", "Hspa1a", "Zfand2a",
  "Ndfip2", "Pam", "Tp53inp1", "Ahi1", "Ube2f", "Hspa4", "Icos", "Chordc1", "Trps1", "Tbc1d4",
  "Rala", "Cd82", "Sema4a", "Pon2", "Acp5", "Ccdc64", "Bhlhe40", "Nampt", "Ahsa1", "Banp",
  "Rhbbd2", "Crem", "Slc7a5", "Cacybp", "Nusap1", "Stip1", "Lrmp", "Pde3b", "Rgs2", "Ccdc141",
  "Snap47", "Dedd2", "Btg3", "Itprip", "Hspa1b", "Galnt2", "Tnfsf9", "Rangap1", "Pdcd1", "Ddx3y",
  "Arid5b", "Dusp10", "Zbtb1", "Samsn1", "Irf4", "Cd2bp2", "Syngr2", "Cdk6", "Mctp2", "Rab27a",
  "Hspd1", "Nfat5", "Batf", "Gzmb", "Neu1", "Syt11", "Cxcr6", "Cnih1", "Fcrl3", "Crtam",
  "Ttn", "Tox", "Morf4l2", "Tigit", "Dctn6", "Mthfd2", "Fut8", "Gata3", "Stat5b", "Akirin2",
  "Whsc1l1", "Stat3", "Zeb2", "Msi2", "B3gnt2", "Eif4e", "Cks2", "Hspe1", "Kpna2", "Ngrn",
  "Ptpn11", "Snap23", "Rasgef1b", "Pfkfb3", "Tnfrsf1b", "Cd27", "Trim13", "Arnt", "Sqstm1", "Mob4",
  "Vps37b", "Arl8b", "Tmem140", "Dnaja1", "Zc3h7a", "GalM", "Stat5a", "Card16", "Amd1", "Rgcc",
  "Golga4", "Sdcbp", "Hnrnpll", "Nr4a1", "Birc3", "Fbxw11", "Tank", "Asxl2", "Eif5", "Fkbp1a",
  "Gspt1", "Jmjd6", "Prdm1", "Laptm4a", "Il21r", "Arpp19", "Fabp5", "Sar1b", "Lyst", "Ezh2",
  "Herpud1", "Traf5", "Anxa5", "Znf331", "Ube2b", "Hbp1", "Sytl3", "Gtf3c1", "Fas", "Sppl2a",
  "Atxn1", "Gga2", "C5orf15", "Tspyl2", "Pmf1", "Rbpj", "Polr3e", "Rapgef1", "Nop58", "Eed",
  "Chst12", "Gtf2b", "Tmx4", "Ccnd2", "Papola", "Fam53b", "Cct4", "Trim59", "Fxr1", "Ppil4",
  "Lag3", "Dnajb6", "Smap2", "GpBp1", "Sertad1", "Pag1", "Trim26", "Tcp1", "Azin1"),

"exhaust_3_feldman" = c("Ccl3", "Epsti1", "Cd38", "Faslg", "Ifi44l", "Gimap6", "Trafd1", "Lgals9", "Cxcr6", "Rab37",
  "Ccr5", "Zbp1", "Samd9l", "Sirpg", "Mx1", "Havcr2", "Acp5", "Ddx60", "Pdcd1", "Sh2d3c",
  "Gpr174", "Rps6ka1", "Gbp5", "Gbp1", "Ptpn6", "S100pbp", "Ifi35", "Oas3", "Snap47", "Gimap4",
  "Parp9", "Ifng", "Sit1", "Pycard", "Rgs3", "Xaf1", "Oas2", "C5orf56", "Gimap5", "Abi3",
  "Snx20", "Vamp5", "Irf2", "Ubash3a", "Parp10", "Gimap7", "Gbp4", "Pvrig", "Cyth4", "Dtx3l",
  "Rhoc", "Sash3", "Ccl4l2", "Ifi6", "Bcas4", "Ikzf3", "Gimap2", "Adora2a", "Arpc5l", "Gyg1",
  "Slfn5", "Chst12", "Apobec3d", "Wars", "Ube2l6", "Tmem140", "Csk", "F2r", "Ctss", "Slamf7",
  "Cxcr3", "Cd27", "Ppp1r18", "Tox", "Ctsc", "Slamf6", "Stat1", "Fut8", "Idh2", "Pced1b",
  "Bst2", "Psmb10", "Stat2", "RnaseT2", "Rbck1", "Sel1l3", "C14orf159", "H2-Aa", "Gzma", "Cd63",
  "Dennd2d", "H2-Ab1", "Prf1", "Cd84", "Tigit", "Ccl4l1", "Plscr1", "Lag3", "Daxx", "Phf11",
  "Igflr1", "Atp6v1e1", "H2-Eb1", "Cmtm3", "Dnajc4", "Lasp1", "H2-Dma", "Nckap1l", "Oasl",
  "Tmem179b", "Ccl4", "Usb1", "Ube2f", "Chmp5", "C19orf66", "Ppm1m", "St8sia4", "Yars",
  "Tbc1d10c", "Drap1", "Pold4", "Trappc1", "Pkn1", "Jakmip1", "Lcp2", "Casp4", "Apol2", "Casp1",
  "Abca2", "H2-DMb1", "Ifi27l2a", "Synrg", "Arhgap30", "Irf7", "Rarres3", "Hmox2", "Lsm2",
  "Gzmh", "Isg15", "Chfr", "Trpv2", "Znhit1", "H2-DPa", "Uba7", "Adam8", "Golim4", "Serpinb1",
  "Atf6b", "Shisa5", "Itgb7", "Tmbim4", "Traf3ip3", "Gpr171", "Traf5", "Arhgef3", "Psmb8",
  "Il2rb", "Apobec3g", "Calcoco2", "Mphosph9", "Dthd1", "Ly6e", "Ppcs", "Capn1", "Gbp2",
  "Pyhin1", "Fkbp1a", "Nudt22", "Ctsd", "Trim14", "Slc25a45", "Klrd1", "Ucp2", "Unc13d",
  "Psmb9", "Gsdmd", "Irf9", "Mpg", "Myo1f", "Slfn12l", "Fermt3", "Mus81", "Apol6", "C17orf62",
  "Fcrl3", "Icam3", "Sp140"
)
)
```

### Computing gene signature with UCell

The function AddModuleScore_UCell() allows operating directly on Seurat objects. UCell scores are calculated from raw counts or normalized data, and returned as metadata columns. The example below defines some simple signatures, and applies them on single-cell data stored in a Seurat object.

```{r fig.width=8, fig.height=5}
# UCELL GENE SIGNATURE SCORE
## Using positive and negative gene sets, GENE+ or GENE-
seurat_obj_T_pseudo <- AddModuleScore_UCell(seurat_obj_T_pseudo, 
                                   features=HumanCancer_gene_signatures, name=NULL)
head(seurat_obj_T_pseudo@meta.data)
```

### Gene signatures
```{r}
gene_signatures_scores_df <- seurat_obj_T_pseudo@meta.data[, c("cell_types", "treatment", names(HumanCancer_gene_signatures))]

gene_signatures_scores_df$cell_types <- factor(gene_signatures_scores_df$cell_types, levels = Tcell_types_order)
gene_signatures_scores_df <- gene_signatures_scores_df %>%
  arrange(cell_types, treatment)

gene_signatures_scores_df$condition <- paste0(gene_signatures_scores_df$cell_types, "__", gene_signatures_scores_df$treatment)
gene_signatures_scores_df$cell_types <- NULL
gene_signatures_scores_df$treatment <- NULL
rownames(gene_signatures_scores_df) <- gene_signatures_scores_df$condition
gene_signatures_scores_df$condition <- NULL
gene_signatures_scores_df
```

```{r fig.height=5, fig.width=15}
pheatmap(t(gene_signatures_scores_df), 
         scale = "row",
         cluster_rows = FALSE, # Cluster rows
         cluster_cols = FALSE, # Cluster columns
         show_rownames = TRUE, # Show row names
         show_colnames = TRUE, # Show column names
         color = colorRampPalette(c("blue", "white", "red"))(100), # Color scale
         border_color = "white",
         main = "Krishna, S., et al. Science (2020) Gene Signatures")
```


# Session Info
```{r}
sessionInfo()
```