---
title: "Final Annotation"
author: "Laura Jiménez Gracia"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.align = "center")
```

# Introduction

## Objective
In this Rmarkdown document, we are going to merge cell type identities annotated by independent cell-lineage analysis into a single one.

# Pre-processing
## Libraries
```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(Seurat)
library(DT)
library(grid)
library(RColorBrewer)
library(scales)
library(gtable)
library(fmsb)
```

## Parameters
Here we will define and load objects and functions that will be used throughout the document.
```{r}
# Paths
path_project_metadata <- here::here("01_cellranger_mapping/data/VEIGAEST_metadata.csv")

path_r_objects <- here::here("03_clustering_annotation/results/R_objects")
path_r_signatures <- here::here("04_GEX_analysis/gene_signatures/results")

# Functions
source(here::here("bin/utils.R"))
`%ni%` <- Negate(`%in%`)

# Parameters
confounder_variables <- c("treatment")
confounder_names <- c("Treatment")

qc_metrics <-  c("nCount_RNA", "nFeature_RNA", "pct_mt", "pct_rb", "S.Score", "G2M.Score")

# Color palette
color_palette <- Polychrome::createPalette(32, c("#fc6060", "#74f774", "#7c7cfc"))
names(color_palette) <- NULL
#Polychrome::swatch(color_palette)

Tcell_types_order <- c("Stem-like CD8 T cells",
  "Effector Memory CD8 T cells",
  "Effector CD8 T cells",
  "IFN-responder CD8 T cells",
  "Exhausted CD8 T cells",
  "Cycling CD8 T cells", 
  "Exhausted CD4 T cells",
  "Tregs", 
  "Cycling Tregs", 
  "Th"
  )

fractionT_color_palette <- c("CD8+ T cells" = "#90CAF9", "CD4+ T cells" = "#D0A9F5")
```

## Load metadata
The data used in this Rmarkdown document comes from **VEIGAEST** dataset.
```{r}
# Load all metadata information
metadata <- read.csv(path_project_metadata)
metadata <- metadata %>% filter(type == "cDNA")
print("Libraries metadata")
DT::datatable(metadata, options = list(scrollX = TRUE))
```

# Load data

```{r}
seurat_obj <- readRDS(paste0(path_r_objects, "/VEIGAEST_clustering_annotation_full_cleaned.rds"))

# Gene signatures
seurat_obj_T_pseudo_metadata <- readRDS(paste0(path_r_signatures, "/VEIGAEST_Tlymphocytes_GeneSignatureScores_metadata.rds"))
```

```{r}
# Subset on conditions of interest
Idents(seurat_obj) <- "treatment"
seurat_obj_cond <- subset(seurat_obj, idents = "BACB", invert = TRUE)
seurat_obj_cond$treatment <- as.character(seurat_obj_cond$treatment)
seurat_obj_cond$treatment <- factor(x = seurat_obj_cond$treatment, levels = c("PBS", "aCD3CD28_BEADS", "BACT"))
seurat_obj_cond
```

```{r}
# Subsetting on T Lymphocytes
seurat_obj_cond_T <- subset(seurat_obj_cond, subset = cell_lineages == "T Lymphocytes")
seurat_obj_cond_T
```



```{r}
Idents(seurat_obj_cond_T) <- "cell_types"
seurat_obj_cond_T <- subset(seurat_obj_cond_T, idents =  c("MAIT", "g/dv2-17 cells", "NKT cells"), invert = TRUE)

# Reorder cell names
seurat_obj_cond_T$cell_types <- as.factor(as.character(seurat_obj_cond_T$cell_types))
seurat_obj_cond_T$cell_types <- factor(x = seurat_obj_cond_T$cell_types, levels = Tcell_types_order)
```


```{r}
# Split T cell fractions
seurat_obj_cond_T$fraction_Tcells <- as.character(seurat_obj_cond_T$cell_types)

seurat_obj_cond_T$fraction_Tcells[seurat_obj_cond_T$cell_types %in% c("Stem-like CD8 T cells","Effector Memory CD8 T cells",
  "Effector CD8 T cells",  "IFN-responder CD8 T cells",  "Exhausted CD8 T cells",  "Cycling CD8 T cells")] <- "CD8+ T cells"

seurat_obj_cond_T$fraction_Tcells[seurat_obj_cond_T$cell_types %in% c("Exhausted CD4 T cells", "Tregs", "Cycling Tregs", "Th")] <- "CD4+ T cells"

seurat_obj_cond_T$fraction_Tcells <- as.factor(seurat_obj_cond_T$fraction_Tcells)

table(seurat_obj_cond_T$fraction_Tcells, seurat_obj_cond_T$treatment)
```

## CD4/CD8 ratio
```{r fig.width=5, fig.height=5}
dittoSeq::dittoBarPlot(seurat_obj_cond_T, 
                       "fraction_Tcells",
                       group.by = "treatment",
                       scale = "percent",
                      retain.factor.levels = TRUE,
                       color.panel = fractionT_color_palette)
```

```{r fig.width=5, fig.height=15}
df <- dittoSeq::dittoBarPlot(seurat_obj_cond_T, 
                       "fraction_Tcells",
                       group.by = "treatment",
                       scale = "percent",
                      retain.factor.levels = TRUE,
                       color.panel = fractionT_color_palette,
                      data.out = TRUE)

# Create the pie chart for each grouping
df$data %>%
  ggplot(aes(x = "", y = percent, fill = label)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  facet_wrap(~ grouping, ncol = 1) +
  scale_fill_manual(values = fractionT_color_palette) +
  theme_void() +
  theme(legend.position = "right",
        legend.text = element_text(size = 12)) +
  labs(fill = "") +
  geom_text(aes(label = count), 
            position = position_stack(vjust = 0.5), 
            color = "black", size = 6)
```
```{r}
genes_of_activation <- c("Il2r2", "Tnfrsf4", "Cd80", "Sos1", "Cd28", "Pdcd1", "Cd247",
                         "Map3k1", "Pak1", "Cd3e", "Cd3d", "Sell", "Cxcr5", "Prkcq", "Pik3r3",
                         "Jun", "Was"
                       )
```

```{r fig.width=8, fig.height=4}
dotplot_cells_activation <- DotPlot(seurat_obj_cond_T,
        features = genes_of_activation,
        group.by = "fraction_Tcells",
        split.by = "treatment",
        scale = TRUE,
        cols = "RdBu",
        dot.scale = 5) +
  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5, face = "italic"))

dotplot_cells_activation
```

## CD8 T cell profiles
```{r}
# Subset on conditions of interest
seurat_obj_cond_TCD8 <- subset(seurat_obj_cond_T, subset = fraction_Tcells == "CD8+ T cells")
seurat_obj_cond_TCD8

# Reorder cell names
seurat_obj_cond_TCD8$cell_types <- as.factor(as.character(seurat_obj_cond_TCD8$cell_types))
seurat_obj_cond_TCD8$cell_types <- factor(x = seurat_obj_cond_TCD8$cell_types, levels = c("Stem-like CD8 T cells",
  "Effector Memory CD8 T cells",
  "Effector CD8 T cells",
  "IFN-responder CD8 T cells",
  "Exhausted CD8 T cells",
  "Cycling CD8 T cells")
  )

table(seurat_obj_cond_TCD8$cell_types, seurat_obj_cond_TCD8$treatment)
```

### Gene Markers
```{r}
genes_of_interest <- c("Tox", "Havcr2", "Pdcd1", # Exhaustion
                       "Gzmk", "Gzmb", "Prf1", "Nkg7", "Ifng", "Tnfrsf9", # Effector molecules
                       "Ccl3", "Ccl4", "Ccl5", "Cxcl13" # Chemokines
                       )
```

```{r fig.width=7, fig.height=8}
dotplot_cells <- DotPlot(seurat_obj_cond_TCD8,
        features = genes_of_interest,
        group.by = "cell_types",
        split.by = "treatment",
        cols = "RdBu",
        dot.scale = 5) +
  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(face = "italic")) + 
  coord_flip()+ 
  scale_colour_gradient2(low="blue", mid="white", high="red")

dotplot_cells

pdf(file = "VEIGAEST_TCD8_byCond_dotplot.pdf", width = 7, height = 8)    
print(dotplot_cells)    
dev.off()   
```

```{r}
genes_of_profile <- c("Sell", "Ccr7", "Lef1", "Bcl2", "Tcf7", # naive survival
                         "Gpr183", "Ltb", "Cxcr3", "Il7r", "Slamf6",  # memory
                         "Isg15", "Ifit1", "Bst2", "Stat1", "Irf7", "Oas3", # IFN response
                         "Lgals3", "Vim", "S100a4", "Anxa2", # cytoskeleton
                         "Lgals1", "Gzmb", "Gzma", "Prf1", "Zeb2", "Klrg1", "Klre1", "Cx3cr1", # effector
                         "Birc5", "Cdk1", "Stmn1", "Mki67", "Top2a" # proliferation
                       )
```

```{r fig.width=14, fig.height=4}
dotplot_cells_profile <- DotPlot(seurat_obj_cond_TCD8,
        features = genes_of_profile,
        group.by = "cell_types",
        #split.by = "treatment",
        scale = TRUE,
        cols = "RdBu",
        dot.scale = 5) +
  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5, face = "italic"))

dotplot_cells_profile
```


### Gene 4BB1

```{r fig.width=9, fig.height=5}
VlnPlot(seurat_obj_cond_TCD8,
        features = "Tnfrsf9",
        group.by = "cell_types",
        split.by = "treatment")
```

### Signature

```{r}
# Subset conditions
seurat_obj_T_pseudo_metadata <- seurat_obj_T_pseudo_metadata %>%
  filter(treatment != "BACB") %>% 
  filter(!cell_types %in% c("MAIT", "g/dv2-17 cells", "NKT cells"))

# Reorder cell names
seurat_obj_T_pseudo_metadata$cell_types <- as.factor(as.character(seurat_obj_T_pseudo_metadata$cell_types))
seurat_obj_T_pseudo_metadata$cell_types <- factor(x = seurat_obj_T_pseudo_metadata$cell_types, levels = Tcell_types_order)

# Split T cell fractions
seurat_obj_T_pseudo_metadata$fraction_Tcells <- as.character(seurat_obj_T_pseudo_metadata$cell_types)
seurat_obj_T_pseudo_metadata$fraction_Tcells[seurat_obj_T_pseudo_metadata$cell_types %in% c("Stem-like CD8 T cells","Effector Memory CD8 T cells",
  "Effector CD8 T cells",  "IFN-responder CD8 T cells",  "Exhausted CD8 T cells",  "Cycling CD8 T cells")] <- "CD8+ T cells"
seurat_obj_T_pseudo_metadata$fraction_Tcells[seurat_obj_T_pseudo_metadata$cell_types %in% c("Exhausted CD4 T cells", "Tregs", "Cycling Tregs", "Th")] <- "CD4+ T cells"
seurat_obj_T_pseudo_metadata$fraction_Tcells <- as.factor(seurat_obj_T_pseudo_metadata$fraction_Tcells)
table(seurat_obj_T_pseudo_metadata$cell_types, seurat_obj_T_pseudo_metadata$treatment)
```

```{r}
seurat_obj_TCD8_pseudo_metadata <- seurat_obj_T_pseudo_metadata %>%
  filter(fraction_Tcells == "CD8+ T cells") %>%
  select(Tc1_phenotype, treatment, cell_types)

seurat_obj_TCD8_pseudo_metadata
```

```{r}
# Prepare data for radar plot
radar_data <- seurat_obj_TCD8_pseudo_metadata %>%
  pivot_wider(names_from = cell_types, values_from = Tc1_phenotype)

# Add the required min/max rows for fmsb
max_values <- rep(0.5, ncol(radar_data) - 1)  # Set maximum values to 1
min_values <- rep(0, ncol(radar_data) - 1)  # Set minimum values to 0

# Combine the min/max rows with actual data
radar_data <- rbind(max_values, min_values, radar_data[,-1])

# Generate the radar plot
radar_chart_colors <- c("red", "blue", "green")  # Customize as needed

radarchart(radar_data,
           axistype = 1,
           pcol = radar_chart_colors, 
           pfcol = adjustcolor(radar_chart_colors, alpha.f = 0.25),
           plwd = 2,
           plty = 1,
           cglcol = "grey", 
           cglty = 1, 
           axislabcol = "grey", 
           vlcex = 1.1,
           title = "Tc1 Phenotype across CD8+ T cells")
           
legend("topright", legend = unique(seurat_obj_T_pseudo_metadata$treatment), 
       col = radar_chart_colors, lty = 1, lwd = 2, bty = "n")
```


## CD4 T cell profiles
```{r}
# Subset on conditions of interest
seurat_obj_cond_TCD4 <- subset(seurat_obj_cond_T, subset = fraction_Tcells == "CD4+ T cells")
seurat_obj_cond_TCD4

# Reorder cell names
seurat_obj_cond_TCD4$cell_types <- as.factor(as.character(seurat_obj_cond_TCD4$cell_types))
seurat_obj_cond_TCD4$cell_types <- factor(x = seurat_obj_cond_TCD4$cell_types, levels = c(
  "Exhausted CD4 T cells",
  "Tregs", 
  "Cycling Tregs", 
  "Th")
  )

table(seurat_obj_cond_TCD4$cell_types, seurat_obj_cond_TCD4$treatment)
```

```{r}
genes_of_interest <- c("Tox", "Havcr2", "Pdcd1", # Exhaustion
                       "Gzmk", "Gzmb", "Prf1", "Nkg7", "Ifng", "Tnfrsf9", # Effector molecules
                       "Ccl3", "Ccl4", "Ccl5", "Cxcl13" # Chemokines
                       )
```

```{r fig.width=6, fig.height=8}
dotplot_cells <- DotPlot(seurat_obj_cond_TCD4,
        features = genes_of_interest,
        group.by = "cell_types",
        split.by = "treatment",
        cols = "RdBu",
        dot.scale = 5) +
  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(face = "italic")) + 
  coord_flip()+ 
  scale_colour_gradient2(low="blue", mid="white", high="red")

dotplot_cells

pdf(file = "VEIGAEST_TCD4_byCond_dotplot.pdf", width = 6, height = 8)    
print(dotplot_cells)    
dev.off()   
```
# Session Info
```{r}
sessionInfo()
```
