---
title: "Final Annotation"
author: "Laura Jiménez Gracia"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.align = "center")
```

# Introduction

## Objective
In this Rmarkdown document, we are going to merge cell type identities annotated by independent cell-lineage analysis into a single one.

# Pre-processing
## Libraries
```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(Seurat)
library(DT)
library(grid)
library(RColorBrewer)
library(scales)
library(gtable)
```

## Parameters
Here we will define and load objects and functions that will be used throughout the document.
```{r}
# Paths
path_project_metadata <- here::here("01_cellranger_mapping/data/VEIGAEST_metadata.csv")

path_r_objects <- here::here("03_clustering_annotation/results/R_objects")
path_r_figs <- here::here("03_clustering_annotation/results/figs")
path_r_tables <- here::here("03_clustering_annotation/results/tables")

# Functions
source(here::here("bin/utils.R"))
`%ni%` <- Negate(`%in%`)

# Parameters
confounder_variables <- c("treatment")
confounder_names <- c("Treatment")

qc_metrics <-  c("nCount_RNA", "nFeature_RNA", "pct_mt", "pct_rb", "S.Score", "G2M.Score")

# Color palette
color_palette <- Polychrome::createPalette(32, c("#fc6060", "#74f774", "#7c7cfc"))
names(color_palette) <- NULL
#Polychrome::swatch(color_palette)
```

## Load metadata
The data used in this Rmarkdown document comes from **VEIGAEST** dataset.
```{r}
# Load all metadata information
metadata <- read.csv(path_project_metadata)
metadata <- metadata %>% filter(type == "cDNA")
print("Libraries metadata")
DT::datatable(metadata, options = list(scrollX = TRUE))
```

# Load data

## Level 1
```{r}
# Load Seurat object -- ANNOT LEVEL 1
seurat_obj <- readRDS(paste0(path_r_objects, "/VEIGAEST_clustering_level1_annotation.rds"))

# Diet Seurat object
seurat_obj <- DietSeurat(
  seurat_obj,
  counts = TRUE,
  data = TRUE,
  scale.data = FALSE,
  assays = "RNA",
  dimreducs = c("pca", "umap", "harmony"),
  graphs = c("RNA_nn", "RNA_snn"))

seurat_obj
```

```{r}
table(seurat_obj$annot_level1)
```

## Level 2
```{r}
annotations_full <- data.frame(annot_level1 = character(),
                               annot_level2 = character()
)

level1_list <- list("Myeloid cells", "T Lymphocytes", "NK cells", "B Lymphocytes", "Neutrophils")
level1_filenames <- c("myeloidcells", "Tlymphocytes", "NKcells", "Blymphocytes", "neutrophils")

# Load & add annotations & merge all together
for (i in 1:length(level1_list)) {
  annot1 <- level1_list[i]
  filename <- level1_filenames[i]
  file_path <- paste0(path_r_objects, "/VEIGAEST_", filename, "_annotation_level2.rds")
  # loading annotations
  annot <- readRDS(file_path)
  # adding annot_level2 labels
  annotations_full <- rbind(annotations_full, annot)
}
annotations_full <- annotations_full["annot_level2"]
```


```{r}
# Add metadata to seurat object
seurat_obj <- AddMetaData(seurat_obj, metadata = annotations_full)

# Replace NA's by level1
seurat_obj$annot_level2[is.na(seurat_obj$annot_level2)] <- seurat_obj$annot_level1[is.na(seurat_obj$annot_level2)]

table(seurat_obj$annot_level2)
length(unique(seurat_obj$annot_level2))
```


## Level 3
```{r}
annotations_full <- data.frame(annot_level1 = factor(),
                               annot_level2 = factor(),
                               annot_level3 = factor()
)

lineage_list <- list("Myeloid cells", "T Lymphocytes", "NK cells", "Neutrophils") 
lineage_filenames <- c("myeloidcells", "Tlymphocytes", "NKcells", "neutrophils")

# Load & add annotations & merge all together
for (i in 1:length(lineage_list)) {
  lineage <- lineage_list[[i]]
  filename <- lineage_filenames[i]
  file_path <- paste0(path_r_objects, "/VEIGAEST_", filename, "_annotation_level3.rds")
  # loading annotations
  annot <- readRDS(file_path)
  # adding annot_level3 labels
  annotations_full <- rbind(annotations_full, annot)
}

annotations_full <- annotations_full["annot_level3"]
```


```{r}
# Add metadata to seurat object
seurat_obj <- AddMetaData(seurat_obj, metadata = annotations_full)

# Replace level3 NA by level2 values
#seurat_obj$annot_level3 <- as.factor(seurat_obj$annot_level3)
seurat_obj$annot_level3[is.na(seurat_obj$annot_level3)] <- seurat_obj$annot_level2[is.na(seurat_obj$annot_level3)]

table(seurat_obj$annot_level3)
length(unique(seurat_obj$annot_level3))
```

## Level 4
```{r}
annotations_full <- data.frame(annot_level1 = factor(),
                               annot_level2 = factor(),
                               annot_level3 = factor(),
                               annot_level4 = factor()
)

lineage_list <- list("Myeloid cells")
lineage_filenames <- c("myeloidcells")

# Load & add annotations & merge all together
for (i in 1:length(lineage_list)) {
  lineage <- lineage_list[[i]]
  filename <- lineage_filenames[i]
  file_path <- paste0(path_r_objects, "/VEIGAEST_", filename, "_annotation_level4.rds")
  # loading annotations
  annot <- readRDS(file_path)
  # adding annot_level3 labels
  annotations_full <- rbind(annotations_full, annot)
}

annotations_full <- annotations_full["annot_level4"]
```


```{r}
# Add metadata to seurat object
seurat_obj <- AddMetaData(seurat_obj, metadata = annotations_full)

# Replace level3 NA by level2 values
#seurat_obj$annot_level3 <- as.factor(seurat_obj$annot_level3)
seurat_obj$annot_level4[is.na(seurat_obj$annot_level4)] <- seurat_obj$annot_level3[is.na(seurat_obj$annot_level4)]

table(seurat_obj$annot_level4)
length(unique(seurat_obj$annot_level4))
```

# Cell naming & grouping
```{r}
seurat_obj
```

## CELL LINEAGES
```{r}
seurat_obj$cell_lineages <- as.character(seurat_obj$annot_level4)

seurat_obj$cell_lineages[seurat_obj$annot_level4 %in% c("Immature B cells", "Mature B cells", "Plasma cells")] <- "B Lymphocytes"
seurat_obj$cell_lineages[seurat_obj$annot_level4 %in% c("N2 TANs","N1 TANs","IFN-response TANs","Mature TANs", "Mast cells")] <- "Granulocytes"

seurat_obj$cell_lineages[seurat_obj$annot_level4 %in% c("g/dv2-17 cells",  "Tregs", "Th", "Effector Memory CD8 T cells",
                                                        "MAIT", "Exhausted CD4 T cells", "Cycling CD8 T cells", 
                                                        "Stem-like CD8 T cells", "Exhausted CD8 T cells", "NKT cells",
                                                        "Cycling Tregs", "IFN-responder CD8 T cells",
                                                        "Effector CD8 T cells")] <- "T Lymphocytes"
seurat_obj$cell_lineages[seurat_obj$annot_level4 %in% c("Arg1+ TAMs", "Ccl8+ TAMs", "Cycling Ccl8+ TAMs",
                                                        "Classical Monocytes")] <- "Monocytes/Macrophages"
seurat_obj$cell_lineages[seurat_obj$annot_level4 %in% c("pDCs","cDC1", "cDC2", "mregDCs", "mo-DCs")] <- "DCs"
seurat_obj$cell_lineages[seurat_obj$annot_level4 %in% c("CD27+CD11b- NK cells", "ILC2", "ILC1", "CD27-CD11b+ NK cells",
                                                        "Cycling NK cells", "IFN-responder NK cells" )] <- "ILCs"


seurat_obj$cell_lineages <- as.factor(seurat_obj$cell_lineages)
table(seurat_obj$cell_lineages)
```

## CELL TYPES
```{r}
seurat_obj$cell_types <- as.character(seurat_obj$annot_level4)
seurat_obj$cell_types <- as.factor(seurat_obj$annot_level4)
table(seurat_obj$cell_types)
length(unique(seurat_obj$cell_types))
```

## Remove annotation slots
```{r}
seurat_obj$annot_level1 <- NULL
seurat_obj$annot_level2 <- NULL
seurat_obj$annot_level3 <- NULL
seurat_obj$annot_level4 <- NULL
```

# Save data

```{r}
# Save metadata
saveRDS(seurat_obj@meta.data, paste0(path_r_objects, "/VEIGAEST_metadata_full.rds"))
write.csv(seurat_obj@meta.data, paste0(path_r_tables, "/VEIGAEST_metadata_full.csv"))
openxlsx::write.xlsx(seurat_obj@meta.data, paste0(path_r_tables, "/VEIGAEST_metadata_full.xlsx"))

# Save annotated Seurat object
saveRDS(seurat_obj, paste0(path_r_objects, "/VEIGAEST_clustering_annotation_full.rds"))
#seurat_obj <- readRDS(paste0(path_r_objects, "/VEIGAEST_clustering_annotation_full.rds"))
```

# Remove uninteresting cells
```{r}
Idents(seurat_obj) <- "cell_lineages"
cells_to_remove <- c("Doublets", "CD45- cells", "LowQuality_cells")
seurat_obj_cleaned <- subset(seurat_obj, idents = cells_to_remove, invert = TRUE)
seurat_obj_cleaned
rm(seurat_obj)
```

## Reorder factors
```{r}
seurat_obj_cleaned$cell_lineages <- as.factor(as.character(seurat_obj_cleaned$cell_lineages))

# Reordering lineages
cell_lineages_list <- c("B Lymphocytes", "DCs", "Monocytes/Macrophages", "T Lymphocytes", "ILCs", "Granulocytes"
                                        )
seurat_obj_cleaned$cell_lineages <- factor(x = seurat_obj_cleaned$cell_lineages,
                                        levels = cell_lineages_list)

table(seurat_obj_cleaned$cell_lineages)
length(unique(seurat_obj_cleaned$cell_lineages))
```


```{r}
seurat_obj_cleaned$cell_types <- as.factor(as.character(seurat_obj_cleaned$cell_types))

# Reordering cells

cell_types_list <- c(# B
                     "Immature B cells",
                     "Mature B cells",
                     "Plasma cells",
                      # DCs
                      "cDC1", "cDC2", "mregDCs", "mo-DCs","pDCs",
                      # Monocytes/Macrophages
                      "Classical Monocytes", "Arg1+ TAMs", "Ccl8+ TAMs", "Cycling Ccl8+ TAMs",
                      # T Lymphocytes
                      "Stem-like CD8 T cells",
                      "Effector CD8 T cells",
                      "Effector Memory CD8 T cells",
                      "IFN-responder CD8 T cells",
                      "Exhausted CD8 T cells",
                      "Cycling CD8 T cells", 
                      "Exhausted CD4 T cells",
                      "Tregs", "Cycling Tregs", 
                      "Th",
                      "MAIT", 
                      "g/dv2-17 cells",
                      "NKT cells",
                      # ILCs
                      "CD27+CD11b- NK cells", "CD27-CD11b+ NK cells", "IFN-responder NK cells", "Cycling NK cells",
                      "ILC1", "ILC2", 
                      "Mature TANs", 
                      "N2 TANs",
                      "N1 TANs",
                      "IFN-response TANs",
                      "Mast cells")
seurat_obj_cleaned$cell_types <- factor(x = seurat_obj_cleaned$cell_types,
                                        levels =  cell_types_list
                                          )

table(seurat_obj_cleaned$cell_types)
length(unique(seurat_obj_cleaned$cell_types))
```

## Dimensionality reduction

```{r}
seurat_obj_cleaned <- seurat_obj_cleaned %>% 
  FindVariableFeatures(nfeatures = 3000) %>% 
  ScaleData() %>% 
  RunPCA() %>% 
  harmony::RunHarmony(group.by.vars = "gem_id") %>%
  RunUMAP(reduction = "harmony", dims = 1:20)
```



# Data exploration

## Lineages
```{r}
# Assign lineages colors
lineages_color_palette <- c("B Lymphocytes" = "#F94144",
                            "DCs" = "#90BE6D",
                            "Monocytes/Macrophages" = "#F9844A",
                            "T Lymphocytes" = "#277DA1",
                            "ILCs" = "#6F4518", 
                            "Granulocytes" = "#132A13"
)
```


```{r fig.width=16, fig.height=12}
# Clustering overview
gg_umap_clustering_cellsubtypes_cleaned <- DimPlot(object = seurat_obj_cleaned,
        group.by = "cell_lineages",
        label = FALSE,
        cols = lineages_color_palette,
        raster = FALSE,
        shuffle = TRUE,
        pt.size = 1
        ) +
  guides(color = guide_legend(override.aes = list(size = 5), ncol = 1)) +
  labs(title="Cell lineages - Cleaned")
gg_umap_clustering_cellsubtypes_cleaned

ggsave(filename = paste0(path_r_figs, "/VEIGAEST_clustering_annotationlineages_full_cleaned_umap.png"),
       plot = gg_umap_clustering_cellsubtypes_cleaned)
```

```{r fig.width=20, fig.height=5}
# Clustering overview
gg_umap_clustering_cellsubtypes_cleaned <- DimPlot(object = seurat_obj_cleaned,
        group.by = "cell_lineages",
        split.by = "treatment",
        label = FALSE,
        cols = lineages_color_palette,
        raster = FALSE,
        pt.size = 0.25,
        ncol = 5
        ) +
  guides(color = guide_legend(override.aes = list(size = 5), ncol = 1)) +
  labs(title="Cell lineages - Cleaned")
gg_umap_clustering_cellsubtypes_cleaned

ggsave(filename = paste0(path_r_figs, "/VEIGAEST_clustering_annotationlineages_full_cleaned_SPLIT_umap.png"),
       plot = gg_umap_clustering_cellsubtypes_cleaned)
```

```{r fig.width=10, fig.height=6}
gg_cell_num <- dittoSeq::dittoBarPlot(seurat_obj_cleaned, 
                            "cell_lineages",
                            group.by = "treatment",
                            scale = "count",
                            legend.show = FALSE,
                            retain.factor.levels = TRUE,
                            color.panel = lineages_color_palette)

gg_cell_prop <- dittoSeq::dittoBarPlot(seurat_obj_cleaned, 
                             "cell_lineages",
                             group.by = "treatment",
                             scale = "percent",
                            retain.factor.levels = TRUE,
                             color.panel = lineages_color_palette)

gg_cell_num + gg_cell_prop
```

##  Cell-types
```{r}

# Assign cell-types colors
cells_color_palette <- c("Immature B cells" = "#ffbaba",
                         "Mature B cells" ="#ff5252",
                         "Plasma cells" = "#a70000",
                        # DCs
                        "cDC1" = "#52B788",
                        "cDC2" = "#1B4332",
                        "mregDCs" = "#4F772D",
                        "mo-DCs" = "#80B918",
                        "pDCs" = "#D8F3DC",
                        # Monocytes/Macrophages
                        "Classical Monocytes" = "#FF6000",
                        "Arg1+ TAMs" = "#F4E409",
                        "Ccl8+ TAMs" = "#583101",
                        "Cycling Ccl8+ TAMs" = "#FFA500",
                        # T Lymphocytes
                        "Stem-like CD8 T cells" = "#89CFF0",
                        "Effector Memory CD8 T cells" = "#1565C0",
                        "Effector CD8 T cells" = "#6A5ACD",
                        "IFN-responder CD8 T cells"= "#003f88",
                        "Exhausted CD8 T cells" = "#2196F3",
                        "Cycling CD8 T cells" = "#BFD7FF",
                        "Exhausted CD4 T cells" = "#9C27B0", 
                        "Tregs" = "#9D4EDD",
                        "Cycling Tregs" = "#7B1FA2", 
                        "Th"= "#C2185B",
                        "MAIT"= "#10002B", 
                        "g/dv2-17 cells"= "#D55AFF",
                        "NKT cells"= "#F8C8DC",
                        # ILCs
                        "CD27+CD11b- NK cells"= "#6F4518",
                        "CD27-CD11b+ NK cells"= "#C36F09", 
                        "IFN-responder NK cells"= "#A47148", 
                        "Cycling NK cells"= "#BC8A5F",
                        "ILC1" = "#eed7a1",
                        "ILC2" = "#fff4f1", 
                         "Mature TANs" = "#D2D5D7", 
                        "N2 TANs" = "#656B69",
                        "N1 TANs" = "#454B47",
                        "IFN-response TANs" = "#020202",
                        "Mast cells" = "#132A13"
                          )
```

```{r fig.width=16, fig.height=12}
# Clustering overview
gg_umap_clustering_cellsubtypes_cleaned <- DimPlot(object = seurat_obj_cleaned,
        group.by = "cell_types",
        label = FALSE,
        cols = cells_color_palette,
        raster = FALSE,
        shuffle = TRUE,
        pt.size = 1
        ) +
  guides(color = guide_legend(override.aes = list(size = 5), ncol = 1)) +
  labs(title="Celltypes - Cleaned")
gg_umap_clustering_cellsubtypes_cleaned

ggsave(filename = paste0(path_r_figs, "/VEIGAEST_clustering_annotationcells_full_cleaned_umap.png"),
       plot = gg_umap_clustering_cellsubtypes_cleaned)
```

```{r fig.width=20, fig.height=8}
gg_cell_num <- dittoSeq::dittoBarPlot(seurat_obj_cleaned, 
                            "cell_types",
                            group.by = "treatment",
                            scale = "count",
                            #legend.show = FALSE,
                            retain.factor.levels = TRUE,
                            color.panel = cells_color_palette)

gg_cell_prop <- dittoSeq::dittoBarPlot(seurat_obj_cleaned, 
                             "cell_types",
                             group.by = "treatment",
                             scale = "percent",
                            retain.factor.levels = TRUE,
                             color.panel = cells_color_palette)

gg_cell_num + gg_cell_prop
```


```{r fig.width=6, fig.height=5}
for (lineage in list("Granulocytes", "DCs", "Monocytes/Macrophages", "T Lymphocytes", "B Lymphocytes","ILCs")) {
  seurat_sub <- subset(seurat_obj_cleaned, subset = cell_lineages == lineage)
  
  # Reorder & susbet
  seurat_sub$cell_types <- as.factor(as.character(seurat_sub$cell_types))
  filtered_cell_types_list <- cell_types_list[cell_types_list %in% unique(seurat_sub$cell_types)]
  seurat_sub$cell_types <- factor(x = seurat_sub$cell_types, levels =  filtered_cell_types_list)
  
  gg_cell_num <- dittoSeq::dittoBarPlot(seurat_sub, 
                            "cell_types",
                            group.by = "treatment",
                            scale = "count",
                            #legend.show = FALSE,
                            retain.factor.levels = TRUE,
                            color.panel = cells_color_palette)
  
    print(gg_cell_num)
}
```


```{r fig.width=6, fig.height=5}
for (lineage in list("Granulocytes", "DCs", "Monocytes/Macrophages", "B Lymphocytes", "T Lymphocytes", "ILCs")) {
  seurat_sub <- subset(seurat_obj_cleaned, subset = cell_lineages == lineage)
  # Reorder & susbet
  seurat_sub$cell_types <- as.factor(as.character(seurat_sub$cell_types))
  filtered_cell_types_list <- cell_types_list[cell_types_list %in% unique(seurat_sub$cell_types)]
  seurat_sub$cell_types <- factor(x = seurat_sub$cell_types, levels =  filtered_cell_types_list)
  
  gg_cell_prop <- dittoSeq::dittoBarPlot(seurat_sub, 
                               "cell_types",
                               group.by = "treatment",
                               scale = "percent",
                              retain.factor.levels = TRUE,
                               color.panel = cells_color_palette)
  
  print(gg_cell_prop)

}
```


# Save data CLEANED
```{r}
# Save metadata
saveRDS(seurat_obj_cleaned@meta.data, paste0(path_r_objects, "/VEIGAEST_metadata_full_cleaned.rds"))
openxlsx::write.xlsx(seurat_obj_cleaned@meta.data, paste0(path_r_tables, "/VEIGAEST_metadata_full_cleaned.xlsx"))
write.csv(seurat_obj_cleaned@meta.data, paste0(path_r_tables, "/VEIGAEST_metadata_full_cleaned.csv"))

# Save annotated Seurat object
saveRDS(seurat_obj_cleaned, paste0(path_r_objects, "/VEIGAEST_clustering_annotation_full_cleaned.rds"))
#seurat_obj_cleaned <- readRDS(paste0(path_r_objects, "/VEIGAEST_clustering_annotation_full_cleaned.rds"))
```

Convert to H5AD format, so downstream analysis can also be run with Scanpy.
```{r}
# Work around to avoid weird problem when saving it as h5ad format
seurat_obj_manual <- CreateSeuratObject(counts = seurat_obj_cleaned@assays$RNA@counts, 
                                        assay = "RNA", 
                                        meta.data = seurat_obj_cleaned@meta.data)
# Keep 2D plotting
seurat_obj_manual@reductions$umap <- seurat_obj_cleaned@reductions$umap

# Modification to ensure it will be readable as AnnData object
# loop over all columns and check if they are numeric
cols_num <- sapply(seurat_obj_manual@meta.data, is.numeric)
# convert non-numeric columns to character type
seurat_obj_manual@meta.data[, !cols_num] <- lapply(seurat_obj_manual@meta.data[, !cols_num], as.character)

intermediate_h5_file_path <- paste0(path_r_objects, "/VEIGAEST_clustering_annotation_full_cleaned.h5Seurat")
# Convert-seurat-to-anndata-object
SeuratDisk::SaveH5Seurat(seurat_obj_manual, filename = intermediate_h5_file_path)

# Then, convert the h5 file into an AnnData object
SeuratDisk::Convert(intermediate_h5_file_path, dest = "h5ad", overwrite = TRUE)

# Remove Seurat h5 intermediate object
base::system2('rm', args = intermediate_h5_file_path)
```


# Gene markers plot

## Cell-types

```{r}
cells_gene_markers <- c(
    # B Lymphocytes
    "Cd79a", "Ms4a1", 
    "Ighd","Cd40", "Ly6a",
    "Jchain",
    ######################################3ADD GENES!!!!  # Cycling    
    "Clec9a", "Xcr1", "Cd24a", #"cDC1"
    "Clec10a", "Mgl2", "Cd209d", #"cDC2"
    "Fscn1", "Ccr7", "Tmem123", #"mregDCs"
    "Il1r2", "Cd209c", #"Mo-DCs" "Itgal", 
    "Siglech", "Ly6d", "Irf8", #"pDCs"
    "Plac8", "Lyz2", "Ms4a4c", #"Classical Monocytes"
    "Cd68", "Apoe", #"M0 Macrophages"
    "Arg1", "Spp1", #"Arg1+ TAM"
    "C1qa","Ccl8", #"Ccl8+ TAM" + "Cycling Ccl8+ TAM",
    # T cells
    "Cd3e", "Cd3d", "Cd8a", "Cd8b1", "Cd4", 
    "Ccr7", "Lef1", "Tcf7", #"Stem-like CD8 T cells"
    "Prdm1", "Tbx21", "Gzmk", "Cx3cr1", #"EFFM CD8 T cells",
    "Xcl1", "Tnfrsf9", "", #"EFF CD8",
    "Isg15", "Ifit3", #"IFN-responder CD8 T cells",
    "Havcr2", "Tox", "Lag3", "Pdcd1",  #"Exhausted CD8 T cells" + "Cycling exhausted CD8+ T cells",  #"Exhausted CD4 T cells",
    "Foxp3","Ctla4", #"Tregs" + "Cycling Tregs", 
    "Icos", "Cd40lg", "Il6ra", #"Th",
    "Klra1",  #"MAIT", 
    "Trdv4", "Il23r", "Il17a",   #"g/dv2-17 cells", 
    #"NKT cells",
    # NK cells
     "Klrb1c", "Klrc1", "Xcl1", "Eomes",
    "Ctla2a", "Itgam", "Prf1", "Gzmb",  #"CD27+CD11b- NK cells", + "Cycling NK cells", #"CD27-CD11b+ NK cells"
    "Ifng", "Ifi206", #"IFN+ NK cells"
   "Itga1", "Rora", "Il7r", # ILC1
     "Gata3","Cxcr6", #"ILC2"
  # Neutrophils
  "S100a8", "S100a9",
  "Retnlg",
    "Cxcl3", "S100a11",
    "Ifit1", "Ifit3", "Cxcl10",
    "Cd200r3", "Gata2",
  "Stmn1",	"Top2a",	"Mki67"	
  )

cells_gene_markers <- cells_gene_markers[!duplicated(cells_gene_markers)]
```


```{r fig.width=18, fig.height=8}
dotplot_cells <- DotPlot(seurat_obj_cleaned,
        features = cells_gene_markers,
        group.by = "cell_types",
        #scale = TRUE,
        #col.max = 3,
        #col.min = 0,
        cols = "RdYlBu",
        dot.scale = 5) +
  #coord_flip() +
  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5, face = "italic"))

dotplot_cells

ggsave(filename = paste0(path_r_figs, "/VEIGAEST_annotationcells_full_genemarkers__dotplot.png"),
       plot = dotplot_cells)
```

# Session Info
```{r}
sessionInfo()
```
