---
title: "Final Annotation"
author: "Laura Jiménez Gracia"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.align = "center")
```

# Introduction

## Objective
In this Rmarkdown document, we are going to merge cell type identities annotated by independent cell-lineage analysis into a single one.

# Pre-processing
## Libraries
```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(Seurat)
library(DT)
library(grid)
library(RColorBrewer)
library(scales)
library(gtable)
```

## Parameters
Here we will define and load objects and functions that will be used throughout the document.
```{r}
# Paths
path_project_metadata <- here::here("01_cellranger_mapping/data/VEIGAEST_metadata.csv")

path_r_objects <- here::here("03_clustering_annotation/results/R_objects")
path_r_figs <- here::here("03_clustering_annotation/results/figs")
path_r_tables <- here::here("03_clustering_annotation/results/tables")

# Functions
source(here::here("bin/utils.R"))
`%ni%` <- Negate(`%in%`)

# Parameters
confounder_variables <- c("treatment")
confounder_names <- c("Treatment")

qc_metrics <-  c("nCount_RNA", "nFeature_RNA", "pct_mt", "pct_rb", "S.Score", "G2M.Score")

# Color palette
color_palette <- Polychrome::createPalette(32, c("#fc6060", "#74f774", "#7c7cfc"))
names(color_palette) <- NULL
#Polychrome::swatch(color_palette)
```

## Load metadata
The data used in this Rmarkdown document comes from **VEIGAEST** dataset.
```{r}
# Load all metadata information
metadata <- read.csv(path_project_metadata)
metadata <- metadata %>% filter(type == "cDNA")
print("Libraries metadata")
DT::datatable(metadata, options = list(scrollX = TRUE))
```

# Load data
```{r}
seurat_obj <- readRDS(paste0(path_r_objects, "/VEIGAEST_clustering_annotation_full_cleaned.rds"))
seurat_obj
```

# Remove uninteresting cells
```{r}
Idents(seurat_obj) <- "treatment"
seurat_obj <- subset(seurat_obj, idents = "BACB", invert = TRUE)
seurat_obj$treatment <- as.factor(as.character(seurat_obj$treatment))
seurat_obj$treatment <- factor(x = seurat_obj$treatment,
                               levels = c("PBS", "aCD3CD28_BEADS", "BACT"))
Idents(seurat_obj) <- "cell_types"
seurat_obj
```


```{r}
table(seurat_obj$treatment)
```

## Dimensionality reduction
```{r}
seurat_obj <- seurat_obj %>% 
  FindVariableFeatures(nfeatures = 3000) %>% 
  ScaleData() %>% 
  RunPCA() %>% 
  harmony::RunHarmony(group.by.vars = "gem_id") %>%
  RunUMAP(reduction = "harmony", dims = 1:30)
```

## Save
```{r}
# Save annotated Seurat object
saveRDS(seurat_obj, paste0(path_r_objects, "/VEIGAEST_clustering_annotation_full_cleaned_BacT.rds"))
#seurat_obj <- readRDS(paste0(path_r_objects, "/VEIGAEST_clustering_annotation_full_cleaned_BacT.rds"))
```


# Data exploration
## Lineages
```{r}
cell_lineages_list <- c("B Lymphocytes", "DCs", "Monocytes/Macrophages", "T Lymphocytes", "ILCs", "Granulocytes")
# Assign lineages colors
lineages_color_palette <- c("B Lymphocytes" = "#F94144",
                            "DCs" = "#90BE6D",
                            "Monocytes/Macrophages" = "#F9844A",
                            "T Lymphocytes" = "#277DA1",
                            "ILCs" = "#6F4518", 
                            "Granulocytes" = "#132A13"
)
```

```{r fig.width=16, fig.height=12}
# Clustering overview
gg_umap_clustering_cellsubtypes_cleaned <- DimPlot(object = seurat_obj,
        group.by = "cell_lineages",
        label = FALSE,
        cols = lineages_color_palette,
        raster = FALSE,
        shuffle = TRUE,
        pt.size = 1
        ) +
  guides(color = guide_legend(override.aes = list(size = 5), ncol = 1)) +
  labs(title="Cell lineages")
gg_umap_clustering_cellsubtypes_cleaned

pdf(file = paste0(path_r_figs, "/gg_umap_clustering_cellsubtypes_cleaned.pdf"), width =16, height =12)    
print(gg_umap_clustering_cellsubtypes_cleaned)    
dev.off()   
```

```{r fig.width=20, fig.height=5}
# Clustering overview
gg_umap_clustering_cellsubtypes_cleaned <- DimPlot(object = seurat_obj,
        group.by = "cell_lineages",
        split.by = "treatment",
        label = FALSE,
        cols = lineages_color_palette,
        raster = FALSE,
        pt.size = 0.25,
        ncol = 5
        ) +
  guides(color = guide_legend(override.aes = list(size = 5), ncol = 1)) +
  labs(title="Cell lineages")
gg_umap_clustering_cellsubtypes_cleaned
```

```{r fig.width=10, fig.height=6}
gg_cell_num <- dittoSeq::dittoBarPlot(seurat_obj, 
                            "cell_lineages",
                            group.by = "treatment",
                            scale = "count",
                            legend.show = FALSE,
                            retain.factor.levels = TRUE,
                            color.panel = lineages_color_palette)

gg_cell_prop <- dittoSeq::dittoBarPlot(seurat_obj, 
                             "cell_lineages",
                             group.by = "treatment",
                             scale = "percent",
                            retain.factor.levels = TRUE,
                             color.panel = lineages_color_palette)

gg_cell_num + gg_cell_prop

pdf(file = paste0(path_r_figs, "/gg_lineages_prop.pdf"), width =6, height =5)    
print(gg_cell_prop)    
dev.off()   
```

##  Cell-types
```{r}
cell_types_list <- c(# B
                     "Immature B cells",
                     "Mature B cells",
                     "Plasma cells",
                      # DCs
                      "cDC1", "cDC2", "mregDCs", "mo-DCs","pDCs",
                      # Monocytes/Macrophages
                      "Classical Monocytes", "Arg1+ TAMs", "Ccl8+ TAMs", "Cycling Ccl8+ TAMs",
                      # T Lymphocytes
                      "Stem-like CD8 T cells",
                      "Effector CD8 T cells",
                      "Effector Memory CD8 T cells",
                      "IFN-responder CD8 T cells",
                      "Exhausted CD8 T cells",
                      "Cycling CD8 T cells", 
                      "Exhausted CD4 T cells",
                      "Tregs", "Cycling Tregs", 
                      "Th",
                      "MAIT", 
                      "g/dv2-17 cells",
                      "NKT cells",
                      # ILCs
                      "CD27+CD11b- NK cells", "CD27-CD11b+ NK cells", "IFN-responder NK cells", "Cycling NK cells",
                      "ILC1", "ILC2", 
                      "Mature TANs", 
                      "N2 TANs",
                      "N1 TANs",
                      "IFN-response TANs",
                      "Mast cells")

# Assign cell-types colors
cells_color_palette <- c("Immature B cells" = "#ffbaba",
                         "Mature B cells" ="#ff5252",
                         "Plasma cells" = "#a70000",
                        # DCs
                        "cDC1" = "#52B788",
                        "cDC2" = "#1B4332",
                        "mregDCs" = "#4F772D",
                        "mo-DCs" = "#80B918",
                        "pDCs" = "#D8F3DC",
                        # Monocytes/Macrophages
                        "Classical Monocytes" = "#FF6000",
                        "Arg1+ TAMs" = "#F4E409",
                        "Ccl8+ TAMs" = "#583101",
                        "Cycling Ccl8+ TAMs" = "#FFA500",
                        # T Lymphocytes
                        "Stem-like CD8 T cells" = "#89CFF0",
                        "Effector Memory CD8 T cells" = "#1565C0",
                        "Effector CD8 T cells" = "#1E90FF",
                        "IFN-responder CD8 T cells"= "#003f88",
                        "Exhausted CD8 T cells" = "#00BFFF",
                        "Cycling CD8 T cells" = "#BFD7FF",
                        "Exhausted CD4 T cells" = "#9C27B0", 
                        "Tregs" = "#9D4EDD",
                        "Cycling Tregs" = "#7B1FA2", 
                        "Th"= "#D55AFF",
                        "MAIT"= "#10002B", 
                        "g/dv2-17 cells"= "#C2185B",
                        "NKT cells"= "#F8C8DC",
                        # ILCs
                        "CD27+CD11b- NK cells"= "#6F4518",
                        "CD27-CD11b+ NK cells"= "#C36F09", 
                        "IFN-responder NK cells"= "#A47148", 
                        "Cycling NK cells"= "#BC8A5F",
                        "ILC1" = "#eed7a1",
                        "ILC2" = "#fff4f1", 
                         "Mature TANs" = "#D2D5D7", 
                        "N2 TANs" = "#656B69",
                        "N1 TANs" = "#454B47",
                        "IFN-response TANs" = "#020202",
                        "Mast cells" = "#132A13"
                          )
```

```{r fig.width=16, fig.height=12}
# Clustering overview
gg_umap_clustering_cellsubtypes_cleaned <- DimPlot(object = seurat_obj,
        group.by = "cell_types",
        label = FALSE,
        cols = cells_color_palette,
        raster = FALSE,
        shuffle = TRUE,
        pt.size = 1
        ) +
  guides(color = guide_legend(override.aes = list(size = 5), ncol = 1)) +
  labs(title="Celltypes")
gg_umap_clustering_cellsubtypes_cleaned

pdf(file = paste0(path_r_figs, "/gg_umap_clustering_cellsubtypes_cleaned.pdf"), width =16, height =12)    
print(gg_umap_clustering_cellsubtypes_cleaned)    
dev.off()   
```

```{r fig.width=20, fig.height=8}
gg_cell_num <- dittoSeq::dittoBarPlot(seurat_obj, 
                            "cell_types",
                            group.by = "treatment",
                            scale = "count",
                            #legend.show = FALSE,
                            retain.factor.levels = TRUE,
                            color.panel = cells_color_palette)

gg_cell_prop <- dittoSeq::dittoBarPlot(seurat_obj, 
                             "cell_types",
                             group.by = "treatment",
                             scale = "percent",
                            retain.factor.levels = TRUE,
                             color.panel = cells_color_palette)

gg_cell_num + gg_cell_prop


pdf(file = paste0(path_r_figs, "/gg_celltypes_prop.pdf"), width =10, height =7)    
print(gg_cell_prop)    
dev.off()   
```


```{r fig.width=6, fig.height=5}
for (lineage in list("Granulocytes", "DCs", "Monocytes/Macrophages", "T Lymphocytes", "B Lymphocytes","ILCs")) {
  seurat_sub <- subset(seurat_obj, subset = cell_lineages == lineage)
  
  # Reorder & susbet
  seurat_sub$cell_types <- as.factor(as.character(seurat_sub$cell_types))
  filtered_cell_types_list <- cell_types_list[cell_types_list %in% unique(seurat_sub$cell_types)]
  seurat_sub$cell_types <- factor(x = seurat_sub$cell_types, levels =  filtered_cell_types_list)
  
  gg_cell_num <- dittoSeq::dittoBarPlot(seurat_sub, 
                            "cell_types",
                            group.by = "treatment",
                            scale = "count",
                            #legend.show = FALSE,
                            retain.factor.levels = TRUE,
                            color.panel = cells_color_palette)
  
    print(gg_cell_num)
  
}
```


```{r fig.width=6, fig.height=5}
for (lineage in list("Granulocytes", "DCs", "Monocytes/Macrophages", "B Lymphocytes", "T Lymphocytes", "ILCs")) {
  seurat_sub <- subset(seurat_obj, subset = cell_lineages == lineage)
  if (lineage == "Monocytes/Macrophages") {
    lineage <- "Mono_Macro"
  }
  # Reorder & susbet
  seurat_sub$cell_types <- as.factor(as.character(seurat_sub$cell_types))
  filtered_cell_types_list <- cell_types_list[cell_types_list %in% unique(seurat_sub$cell_types)]
  seurat_sub$cell_types <- factor(x = seurat_sub$cell_types, levels =  filtered_cell_types_list)
  
  gg_cell_prop <- dittoSeq::dittoBarPlot(seurat_sub, 
                               "cell_types",
                               group.by = "treatment",
                               scale = "percent",
                              retain.factor.levels = TRUE,
                               color.panel = cells_color_palette)
  
  print(gg_cell_prop)
  
  pdf(file = paste0(path_r_figs, "/gg_", lineage, "_celltypes_prop.pdf"), width =6, height =5) 
  print(gg_cell_prop)    
  dev.off()   
}
```


# Gene markers plot

## Cell-types

```{r}
cells_gene_markers <- c(
    # B Lymphocytes
    "Cd79a", "Ms4a1", 
    "Ighd","Cd40", "Ly6a",
    "Jchain",
    ######################################3ADD GENES!!!!  # Cycling    
    "Clec9a", "Xcr1", "Cd24a", #"cDC1"
    "Clec10a", "Mgl2", "Cd209d", #"cDC2"
    "Fscn1", "Ccr7", "Tmem123", #"mregDCs"
    "Cd209c", "Il1r2", #"Mo-DCs" "Itgal", 
    "Siglech", "Ly6d", "Irf8", #"pDCs"
    "Lyz2", "Plac8","Ms4a4c", #"Classical Monocytes"
    "Cd68", "Apoe",
    "Arg1", "Spp1", #"Arg1+ TAM"
    "C1qa","Ccl8", #"Ccl8+ TAM" + "Cycling Ccl8+ TAM",
    # T cells
    "Cd3e", "Cd3d", "Cd8a", "Cd8b1", "Cd4", 
   "Lef1", "Ccr7",  "Tcf7", #"Stem-like CD8 T cells"
    "Xcl1", "Tnfrsf9", "Gpr183", #"EFF CD8",
    "Tbx21", "Gzmk", "Cxcr3", #"EFFM CD8 T cells",
    "Isg15", "Ifit3", #"IFN-responder CD8 T cells",
    "Tox", "Havcr2", "Pdcd1","Lag3",  #"Exhausted CD8 T cells"  #"Exhausted CD4 T cells",
    "Foxp3","Ctla4", #"Tregs" + "Cycling Tregs", 
    "Icos", "Cd40lg", #"Th",
    "Klra1",  #"MAIT", 
    "Trdv4", "Il23r", "Il17a",   #"g/dv2-17 cells", 
    "Nkg7",#"NKT cells",
    # NK cells
     "Klrb1c", "Klrc1", "Xcl1", "Eomes",
    "Ctla2a", "Itgam", "Prf1", "Gzmb",  #"CD27+CD11b- NK cells", + "Cycling NK cells", #"CD27-CD11b+ NK cells"
    "Ifng", "Ifi206", #"IFN+ NK cells"
   "Itga1", "Rora", "Il7r", # ILC1
     "Gata3","Cxcr6", #"ILC2"
  # Neutrophils
  "S100a8", "S100a9",
  "Retnlg",
    "Cxcl3", "S100a11",
    "Ifit1", "Ifit3", "Cxcl10",
    "Cd200r3", "Gata2",
  "Stmn1",	"Top2a",	"Mki67"	
  )

cells_gene_markers <- cells_gene_markers[!duplicated(cells_gene_markers)]
```


```{r fig.width=18, fig.height=8}
seurat_obj$cell_types <- factor(x = seurat_obj$cell_types, levels = rev(cell_types_list))

dotplot_cells <- DotPlot(seurat_obj,
        features = cells_gene_markers,
        group.by = "cell_types",
        #scale = TRUE,
        #col.max = 3,
        #col.min = 0,
        cols = "RdYlBu",
        dot.scale = 5) +
  #coord_flip() +
  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5, face = "italic"))

dotplot_cells

pdf(file = paste0(path_r_figs, "/dotplot_genemarkers_cells.pdf"), width =18, height =8)
print(dotplot_cells)    
dev.off()   
```

# Dotplots populations by lineage
## T
```{r fig.width=7, fig.height=5}
Idents(seurat_obj) <- "cell_lineages"

seurat_obj_T <- subset(seurat_obj, idents = "T Lymphocytes")

T_cell_names <- c("Stem-like CD8 T cells",
                  "Effector CD8 T cells",
                  "Effector Memory CD8 T cells",
                  "IFN-responder CD8 T cells",
                  "Exhausted CD8 T cells",
                  "Cycling CD8 T cells", 
                  "Exhausted CD4 T cells",
                  "Tregs", "Cycling Tregs", 
                  "Th",
                  "MAIT", 
                  "g/dv2-17 cells",
                  "NKT cells")

seurat_obj_T$cell_types <- factor(x = seurat_obj_T$cell_types, levels = rev(T_cell_names))

T_gene_markers <- c("Lef1", 
                    "Grp183", "Tnfrsf9",
                    "Gpr183",
                    "Isg15",
                    "Gzmb", "Prf1", "Tox", 
                     "Foxp3",
                    "Cd40lg", "Klra1",
                    "Il23r",
                    "Nkg7",
                    "Mki67")

dotplot_T_cells <- DotPlot(seurat_obj_T,
        features = T_gene_markers,
        group.by = "cell_types",
        scale = TRUE,
        cols = "RdYlBu",
        dot.scale = 4) +
  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5, face = "italic")
        )

dotplot_T_cells

pdf(file = paste0(path_r_figs, "/dotplot_genemarkers_TLymphocytes.pdf"), width =7, height =5)
print(dotplot_T_cells)    
dev.off() 
```

## Mono/Macro
```{r fig.width=5, fig.height=3}
Idents(seurat_obj) <- "cell_lineages"

seurat_obj_MoMa <- subset(seurat_obj, idents = "Monocytes/Macrophages")

MoMa_cell_names <- c("Classical Monocytes", "Arg1+ TAMs", "Ccl8+ TAMs", "Cycling Ccl8+ TAMs")

seurat_obj_MoMa$cell_types <- factor(x = seurat_obj_MoMa$cell_types, levels = rev(MoMa_cell_names))

MoMa_gene_markers <- c("Lyz2", "Arg1", "Ccl8", "Mki67")

dotplot_MoMa_cells <- DotPlot(seurat_obj_MoMa,
        features = MoMa_gene_markers,
        group.by = "cell_types",
        scale = TRUE,
        cols = "RdYlBu",
        dot.scale = 4) +
  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5, face = "italic")
        )

dotplot_MoMa_cells

pdf(file = paste0(path_r_figs, "/dotplot_genemarkers_MonoMacro.pdf"), width =5, height =3)
print(dotplot_MoMa_cells)    
dev.off() 
```

## DC
```{r fig.width=5, fig.height=3}
Idents(seurat_obj) <- "cell_lineages"

seurat_obj_DC <- subset(seurat_obj, idents = "DCs")

DC_cell_names <- c("cDC1", "cDC2", "mregDCs", "mo-DCs","pDCs")

seurat_obj_DC$cell_types <- factor(x = seurat_obj_DC$cell_types, levels = rev(DC_cell_names))

DC_gene_markers <- c("Clec9a", "Clec10a", "Fscn1","Cd209c", "Siglech")

dotplot_DC_cells <- DotPlot(seurat_obj_DC,
        features = DC_gene_markers,
        group.by = "cell_types",
        scale = TRUE,
        cols = "RdYlBu",
        dot.scale = 4) +
  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5, face = "italic")
        )

dotplot_DC_cells

pdf(file = paste0(path_r_figs, "/dotplot_genemarkers_DC.pdf"), width =5, height =3)
print(dotplot_DC_cells)    
dev.off() 
```

## Granulocytes
```{r fig.width=5, fig.height=3}
Idents(seurat_obj) <- "cell_lineages"

seurat_obj_Granu <- subset(seurat_obj, idents = "Granulocytes")

Granu_cell_names <- c("Mature TANs", 
                      "N2 TANs",
                      "N1 TANs",
                      "IFN-response TANs",
                      "Mast cells")

seurat_obj_Granu$cell_types <- factor(x = seurat_obj_Granu$cell_types, levels = rev(Granu_cell_names))

Granu_gene_markers <- c("Retnlg",
    "Cxcl3", "S100a11",
    "Ifit1", 
    "Cd200r3")

dotplot_Granu_cells <- DotPlot(seurat_obj_Granu,
        features = Granu_gene_markers,
        group.by = "cell_types",
        scale = TRUE,
        cols = "RdYlBu",
        dot.scale = 4) +
  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = 0.5, face = "italic")
        )

dotplot_Granu_cells

pdf(file = paste0(path_r_figs, "/dotplot_genemarkers_Granulocytes.pdf"), width =5, height =3)
print(dotplot_Granu_cells)    
dev.off() 
```

# Session Info
```{r}
sessionInfo()
```
